---
title: "Analysis of Top 50 Best Selling Books on the basis of Genre (YEAR 2009 TO 2019)"
author: "Isha Dubey"
date: "3/25/2022"
output: html_document
---
```{r include=FALSE,echo=FALSE,error=FALSE,}
#importing packages and loading them

#install.packages("tidyverse")
#install.packages("lubridate")
#install.packages("dplyr")
#install.packages("ggplot2")
#install.packages("sqldf")
#install.packages("tidyr")

library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
library(sqldf)
library(rmarkdown)
library(tidyr)

getwd()
setwd("C:\\Users\\ishad\\Documents\\Isha\\RCaseStudy\\RCaseStudy3\\AmazonBestSellingBooks")

#reading book sellers csv into a dataframe
bookSeller <- read.csv("bestsellers with categories.csv")
#View(bookSeller)

#colnames(bookSeller)
#str(bookSeller)

#adding columns to bookSeller
#Year1 to use it for graph
bookSeller <- mutate(bookSeller,Year1=as.character(bookSeller$Year))
#user_rating to use it in sql queries
bookSeller <- mutate(bookSeller,user_rating=bookSeller$User.Rating)


#-------------------------------Analyizing Data ------------------
#FInding max user rating
max_rating <- max(bookSeller$User.Rating)
print(max_rating)
#FInding min user rating
min_rating <- min(bookSeller$User.Rating)
print(min_rating)
#Storing various rating from 2009 to 2019 for the books
user_rating_values <- sqldf("select distinct user_rating from bookSeller")
#print(user_rating_values)

user_rating_values <- sort(user_rating_values$user_rating,decreasing=TRUE)
#print(user_rating_values)
```
## =================================================

### **NUMBER OF BOOKS BASED ON GENRE FROM YEAR 2009 TO 2019**


```{r echo=FALSE,error=FALSE}
#===============================================================
#counting Number of books on the basis of Genre out of top 50 best selling book through year 2009 to 2019
Genre_Count_Book <- count(bookSeller,Genre)
colnames(Genre_Count_Book)<- c("Genre","No_of_Books")
#View(Genre_Count_Book)
ggplot(Genre_Count_Book)+
  geom_col(mapping = aes(x=Genre,y = No_of_Books,fill=No_of_Books))+
  labs(title = "Count of books on the basis of Genre",
       subtitle = "Year 2009 to 2019",x="Genre",y="Number of Books" ,caption = "( Graph 1 )")

#===================================================================
```

##### **CONCLUSION**
- Number of Fiction Books are less as compared to Non-Fiction Books.

## =================================================

### **COUNTING NUMBER OF BOOKS EACH YEAR FROM 2009 TO 2019 BASED ON GENRE** 

```{r echo=FALSE}
#Observing Number of books on the basis of Genre per year
Count_Genre_PerYear <- count(bookSeller,Genre,Year1) 
colnames(Count_Genre_PerYear)<- c("Genre","Year1","No_of_Books")
#View(Count_Genre_PerYear)

ggplot(Count_Genre_PerYear)+
  geom_col(mapping = aes(x=Year1,y=No_of_Books,fill=Genre))+
  labs(title = "Number of books on the basis of Genre per year",
       subtitle = "Year 2009 to 2019", x="Year ->",y="Number of Books ->",caption = "( Graph 2 )")

#===================================================================
```

##### **CONCLUSION**
- Number of Non-fiction books is more as compared to the fiction books except year 2014.


## =================================================

### **FINDING TOP TEN AUTHORS FROM YEAR 2009 TO 2019 AND OBSERVING THEM BASED ON GENRE**

```{r echo=FALSE}
TopTenPopularAuthor09To19 <- sqldf("select Author,count(Author)booksWritten,Genre
                                 from bookSeller 
                                 group by Author
                                 order by booksWritten desc 
                                 limit 10")

#View(TopTenPopularAuthor09To19)

ggplot(TopTenPopularAuthor09To19)+
  geom_col(mapping = aes(x=Author,y=booksWritten,fill=Genre))+
  labs(title="Top 10 Authors on basis of Count of Books Written",
       subtitle = "From Year 2009 To 2019",y="Count of Books Written->", x="Author Name ->",caption = "( Graph 3 )")+
  theme(axis.text.x = element_text(angle=90))

#===================================================================
```

##### **CONCLUSION**
- Number of Fiction writer are more as compared to Non-fiction writers


## =================================================

### **FINDING FAVORITE SERIES FROM YEAR 2009 TO 2019**

```{r echo=FALSE ,error=FALSE,warning=FALSE}

#FINDING FAVORITE SERIES FROM YEAR 2009 TO 2019
bookSeller1 <- separate(bookSeller,Name,c("Name2"),":",convert = FALSE)
#View(bookSeller1)

Longest_series <- count(bookSeller1,Name2,Genre,user_rating,sort = TRUE)
colnames(Longest_series)<-c("Name","Genre","user_rating","No_of_Books")
#View(Longest_series)

FavoriteLongestSeries <-
  sqldf("select * from Longest_series 
        where user_rating=4.9 AND No_of_Books>4")
#View(FavoriteLongestSeries)

ggplot(FavoriteLongestSeries)+
  geom_col(mapping = aes(x=Name,y=No_of_Books,fill=Genre))+
  labs(title="Longest favorite Book Series",subtitle = "From Year     2009 to 2019", x="Name of the book series ->", y="Number of    Books ->" ,caption = "( Graph 4 )")+
  theme(axis.text.x=element_text(angle=90))
```

##### **CONCLUSION**
- Longest and Favorite books series are fictional.

## =================================================

### **FINDING THE BEST BOOK OF THE YEAR FROM YEAR 2009 TO 2019**


```{r echo=FALSE,,warning=FALSE,fig.width=8,fig.height=10}

#===================================================================

#Finding the best book the Year from 2009 to 2019
BestBook2009To2019 <- sqldf("select * from bookSeller 
           where user_rating>4.7
           group by Year1
           order by Year1")
#View(BestBook2009To2019)


ggplot(BestBook2009To2019)+
  geom_col(mapping = aes(x=Reviews,y=Name,fill=Year1,color=Genre))+
  labs(title="Best book of the year (User Rating > 4.7)",subtitle = "From year 2009 to 2019", x="Year ->",y="Name of the Book ->" ,caption = "( Graph 5    )")+
  theme(axis.text.x=element_text(angle=90),legend.position = "bottom",legend.box = "vertical")+
  scale_y_discrete(labels=label_wrap_gen(width = 25,multi_line = TRUE))+
  annotate("text", x =22000, y = 6,
                     label = "Non-Fiction")
#===================================================================


```

##### **CONCLUSION**
- All the years have Fiction Books as the Best book of the year except the Book "Jesus Calling:Enjoying Peace in his Presence(With Scriptures References)" in Year 2011 which is Non-Fiction.


## =================================================

#### **FINAL CONCLUSION FROM ALL THE GRAPHS**
- Though the number of non-fiction books are more in top 50 best selling books, but the fiction books leads the non-fiction books.

## =================================================


\