---
title: "Smart Device Usage Analysis For Bellabeat"
author: "Isha Dubey"
date: "3/21/2022"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r include=FALSE,echo=FALSE,error=FALSE}
#importing packages and loading them
#install.packages("tidyverse")
#install.packages("lubridate")
#install.packages("writexl")
#install.packages("xlsx")
#install.packages("dplyr")
#install.packages("xlsx")
#install.packages("ggplot2")
#install.packages("sqldf")
#install.packages("rmarkdown")

library(tidyverse)
library(readxl)
library(dplyr)
library(lubridate)
library(writexl)
library(dplyr)
library(ggplot2)
library(sqldf)
library(rmarkdown)
library(tinytex)
#============================importing files=======================================
getwd()
setwd("C:\\Users\\ishad\\Documents\\Isha\\RCaseStudy\\RCaseStudy2\\FitabaseData")
#------------------------------------------------------------------------------------
dailyActivity <- 
  read.csv("C:\\Users\\ishad\\Documents\\Isha\\RCaseStudy\\RCaseStudy2\\FitabaseData\\dailyActivity_merged.csv")
#View(dailyActivity)

sleepDay <-
  read.csv("C:\\Users\\ishad\\Documents\\Isha\\RCaseStudy\\RCaseStudy2\\FitabaseData\\sleepDay_merged.csv")
#View(sleepDay)

heartRateSec <-
  read.csv("C:\\Users\\ishad\\Documents\\Isha\\RCaseStudy\\RCaseStudy2\\FitabaseData\\heartrate_seconds_merged.csv")
#View(heartRateSec)

weightLog <- 
  read.csv("C:\\Users\\ishad\\Documents\\Isha\\RCaseStudy\\RCaseStudy2\\FitabaseData\\weightLogInfo_merged.csv")
#View(weightLog)

#================================ORGANISING DATA===================================
#STEP 1
# adding columns AvgDistance and AvgActiveM into dailyActivity data frame 
# adding column DateOfActivity_YMD, extracted from the ActivityDate 
actDate <- parse_date_time(dailyActivity$ActivityDate, "m/d/Y")

dailyActivity2 <- 
mutate(dailyActivity,
       AvgDistance=(VeryActiveDistance+ModeratelyActiveDistance+
                    LightActiveDistance+SedentaryActiveDistance)/4,
       AvgActiveMin=(VeryActiveMinutes+FairlyActiveMinutes+
                      LightlyActiveMinutes)/3,
       DateOfActivity_YMD = as.Date(actDate,format="%Y/%m/%d")
      )

# changing type of Id to character from numeric (to use for color attribute in ggplot)
class(dailyActivity2$Id)="character"

str(dailyActivity2)

#STEP 2
#adding columns sleep_Date and Sleep_Time to sleepDay data frame extracted from sleepDay column 
SlpTime = parse_date_time(sleepDay$SleepDay, "m/d/Y I:M:S p")
sleepDay2 <- 
mutate(sleepDay,sleep_Date = as.Date(SlpTime,format = "%Y/%m/%d" ),
Sleep_Time =format(SlpTime,format = "%H:%M:%S") )


#changing type of Id to character from numeric (to use for color attribute in ggplot)
class(sleepDay2$Id)="character"

str(sleepDay2)

#STEP 3
# MERGING DATA FRAMES OF dailyActivity AND sleepDay to see relation between Activity and Sleep
MergeActivitySleep <- 
sqldf("select d.Id, d.TotalSteps,d.TotalDistance,
      d.TrackerDistance,d.LoggedActivitiesDistance,
      d.SedentaryMinutes,d.Calories,d.AvgDistance,
      d.AvgActiveMin,d.DateOfActivity_YMD,s.sleep_Date,
      s.Sleep_Time,s.TotalSleepRecords,s.TotalMinutesAsleep,
      s.TotalTimeInBed
      from dailyActivity2 d
      inner join sleepDay2 s
      on d.Id = s.Id
      where d.DateOfActivity_YMD == s.sleep_Date
      ")
#View(MergeActivitySleep)

#STEP 4
# adding columns to heartRateSec data frame heartRateDate_YMD and heartRateTime from column Time 
SleepTime = parse_date_time(heartRateSec$Time, "m/d/Y I:M:S p")
heartRateSec1 <- mutate(heartRateSec, heartRateDate_YMD =  as.Date(SleepTime,format = "%Y/%m/%d"),heartRateTime =format(SleepTime,format = "%H:%M:%S")  )

# changing data type of id to character from numeric
class(heartRateSec1$Id)="character"

str(heartRateSec1)

#STEP 5
# CERATING heart rate summary data frame by finding average/max/min heartrates for each user
heartRateSmry <- 
  sqldf("select Id,heartRateDate_YMD,count(Id) as HeartRateReadingCount, 
      avg(Value) as AverageHeartRate,
      max(Value) as MaxHeartRate,min(Value) as MinHeartRate 
      from heartRateSec1 
      group by Id, heartRateDate_YMD")
#View(heartRateSmry)
colnames(heartRateSmry)
#colnames(MergeActivitySleep)

#STEP 6
# MERGING DATA FRAME FOR "dailyActivity and sleep" with heartratesec data frame
MergeActivitySleepHeartRate <-
  sqldf("select MergeActivitySleep.*,
        m.heartRateDate_YMD,m.HeartRateReadingCount,
        m.AverageHeartRate,m.MaxHeartRate,m.MinHeartRate
        from MergeActivitySleep
        inner join heartRateSmry m
        on MergeActivitySleep.Id = m.Id
        where MergeActivitySleep.DateOfActivity_YMD == m.heartRateDate_YMD
      ")
#View(MergeActivitySleepHeartRate)
str(MergeActivitySleepHeartRate)

#STEP 7
#changing data type of id and log id to character from numeric 
class(weightLog$Id)="character"
class(weightLog$LogId)="character"

#adding columns to weightLog data frame
wtLogTime = parse_date_time(weightLog$Date, "m/d/Y I:M:S p")
weightLog2 <- 
  mutate(weightLog,WeightLogDate_YMD = as.Date(wtLogTime,format = "%Y/%m/%d" ),
         WeightLogTime =format(wtLogTime,format = "%H:%M:%S") )
#View(weightLog2)
colnames(weightLog2)

#MERGING DATA FROM MergeActivitySleepHeartRate AND weightLog2
MergeActivitySleepHeartRateWeight <-
  sqldf("select MergeActivitySleepHeartRate.*,
    w.WeightKg,w.WeightPounds,w.BMI,w.IsManualReport,w.LogId,
    w.WeightLogDate_YMD,w.WeightLogTime
    from MergeActivitySleepHeartRate
    inner join weightLog2 w
    on MergeActivitySleepHeartRate.Id = w.Id
    where MergeActivitySleepHeartRate.DateOfActivity_YMD == w.WeightLogDate_YMD")

#View(MergeActivitySleepHeartRateWeight)
str(MergeActivitySleepHeartRateWeight)

```

## -----------------------------------------------------------------------------------------

### RELATION B/W THE STEPS AND DISTANCE TRACKED BY THE DEVICE

```{r echo=FALSE}
#======================== CREATING VISUALIZATIONS ===================================

#RELATION BETWEEN TOTOL STEPS AND DISTANCE INTENSITIES
  #1 relation between steps and TrackerDistance
ggplot(dailyActivity2)+
  geom_point(mapping = aes(x=TotalSteps,y=TrackerDistance,color=Id))+
  labs(title="Relation b/w Total Steps Taken in a day and Tracked Distance by the device",x="Total Steps ->", y="Distance Tracked by the device ->",caption = "(Graph 1)")
  #2 relation between steps and Total Distance
ggplot(dailyActivity2)+
  geom_point(mapping = aes(x=TotalSteps,y=TotalDistance,color=Id))+
  labs(title="Relation b/w Total Steps taken in a day and Total Distance convered in a day", x="TotalSteps", y="TotalDistance",caption = "(Graph 2)")
  #3 relation between steps and average distance 
ggplot(dailyActivity2)+
  geom_point(mapping = aes(x=TotalSteps,y=AvgDistance,color=Id))+
  labs(title="Relation b/w Total Steps taken in a day and Average Distance convered in a day",x="TotalSteps ->",y="AvgDistance->",caption="(Graph 3)")


##====================================================================================

```


#### **CONCLUSION FROM GRAPH 1, GRAPH 2 AND GRAPH 3**
- The Relationship between steps and distance is linear.
- Distance is increasing with the steps.
- The Average disance is the average of the following distances : 
  * Very Active
  * Moderately Active
  * Lightly Active


## -----------------------------------------------------------------------------------------

### RELATION BETWEEN THE STEPS TAKEN AND CALORIES BURNT
  
```{r echo=FALSE}

#RELATION BETWEEN TOTAL STEPS AND CALORIES
ggplot(dailyActivity2)+
  geom_point(mapping = aes(x=TotalSteps,y=Calories,color=Id))+
  labs(title = "Relation b/w TotalSteps and Calories burnt",caption = "(Graph 4)")
##====================================================================================
```

#### **CONCLUSION FROM GRAPH 4**
- The Relationship between steps and calories shows the increasing trend.
- More calories are burnt as the steps increase.

## -----------------------------------------------------------------------------------------

### RELATION BETWEEN THE STEPS TAKEN AND ACTIVITY MINUTES

```{r echo=FALSE}
#RELATION BETWEEN TOTAL STEPS AND ACTIVITY MINUTES
  #1 Active minutes
ggplot(dailyActivity2)+
  geom_point(mapping = aes(x=TotalSteps,y=AvgActiveMin,color=Id))+
  labs(title = "Relation b/w TotalSteps and Average Activity Minutes",x="Total Steps taken ->",y="Average Activity Minutes ->",caption = "(Graph 5)")
  #2 Sedentary minutes
ggplot(dailyActivity2)+
  geom_point(mapping = aes(x=TotalSteps,y=SedentaryMinutes,color=Id))+
  labs(title = "Relation b/w TotalSteps and Sedentary Minutes",caption = "(Graph 6)")
#====================================================================================
```

#### **CONCLUSION FROM GRAPH 5**
- The Relationship between steps and Average Activity minutes shows the increasing trend.
- More steps are counted as the activity minutes are increased.
- Average Activity minutes is the average of the following:
  * VeryActiveMinutes 
  * FairlyActiveMinutes 
  * LightlyActiveMinutes

#### **CONCLUSION FROM GRAPH 6**
- The Steps and Sedentary Minutes seems to have decreasing trend relation.

## -----------------------------------------------------------------------------------------
### RELATION BETWEEN THE AVERAGE ACTIVITY MINUTES AND AVERAGE DISTANCE INTENSITY

```{r echo=FALSE}
#RELATION BETWEEN AVERAGE ACTIVITY MINUTES AND AVERAGE DISTANCE INTENSITY
ggplot(dailyActivity2)+
  geom_point(mapping = aes(x=AvgActiveMin,y=AvgDistance,color=Id))+
  labs(title = "Relation b/w Average Activity Minutes and Average Distance",caption = "(Graph 7)")
```

#### **CONCLUSION FROM GRAPH 7**
- The Graph shows the increasing trend.
- It concludes that the more is the distance covered more is the activity done.

## -----------------------------------------------------------------------------------------
### RELATION BETWEEN THE CALORIES BURNT AND SLEEP

```{r echo=FALSE}
#RELATION BETWEEN CALORIES AND SLEEP
ggplot(MergeActivitySleep)+
  geom_point(mapping = aes(x=Calories,y=TotalMinutesAsleep,color=Id))+
  labs(title = "Relation b/w Calories burnt and Total sleeping minutes",caption = "(Graph 8)")
#========================================================================================

```

#### **CONCLUSION FROM GRAPH 8**
- The Graph shows the scattered trend.
- To identify the actual trend individual user has to be identified.

## -----------------------------------------------------------------------------------------
### RELATION BETWEEN THE TOTAL STEPS AND SLEEP

```{r echo=FALSE}
#RELATION BETWEEN TOTAL STEPS AND SLEEP
ggplot(MergeActivitySleep)+
  geom_point(mapping = aes(x=TotalSteps,y=TotalMinutesAsleep,color=Id))+
  labs(title = "Relation b/w TotalSteps and Total sleeping minutes",caption = "(Graph 9)")

#RELATION BETWEEN AVERAGE DISTANCE INTENSITIES  AND SLEEP
ggplot(MergeActivitySleep)+
  geom_point(mapping = aes(x=AvgDistance,y=TotalMinutesAsleep,color=Id))+
  labs(title = "Relation b/w Average distance intensities and Total sleeping minutes", caption = "(Graph 10)")

#RELATION BETWEEN AVERAGE ACTIVITY MINUTES AND SLEEP
ggplot(MergeActivitySleep)+
  geom_point(mapping = aes(x=AvgActiveMin,y=TotalMinutesAsleep,color=Id))+
  labs(title = "Relation b/w Average active minutes and Total sleeping minutes", caption = "(Graph 11)")

```

#### **CONCLUSION FROM GRAPH 9, GRAPH 10 AND GRAPH 11**
- The Relationship between Distance and sleep is scattered.
- Trend for an individual may be identified one at a time to find the relation.
- The Average disance is the average of the following distances : 
  * Very Active
  * Moderately Active
  * Lightly Active
- Average Activity minutes is the average of the following:
  * VeryActiveMinutes 
  * FairlyActiveMinutes 
  * LightlyActiveMinutes

## -----------------------------------------------------------------------------------------
### RELATION BETWEEN THE TOTAL STEPS AND HEARTRATE

```{r echo=FALSE}
#RELATION BETWEEN STEPS AND HEARTRATE
  #1  Average Heart Rate
ggplot(MergeActivitySleepHeartRate)+
  geom_point(mapping = aes(x=TotalSteps,y=AverageHeartRate,color=Id))+
  labs(title = "Relation b/w Total Steps and Average Heart Rate",caption = "(Graph 12)")
  #2  Maximum Heart Rate
ggplot(MergeActivitySleepHeartRate)+
  geom_point(mapping = aes(x=TotalSteps,y=MaxHeartRate,color=Id))+
  labs(title = "Relation b/w Total Steps and Maximum Heart Rate for each user ", caption = "(Graph 13)")
  #3  Minimum Heart Rate
ggplot(MergeActivitySleepHeartRate)+
  geom_point(mapping = aes(x=TotalSteps,y=MinHeartRate,color=Id))+
  labs(title = "Relation b/w Total Steps and Minimum Heart Rate for each user",caption = "(Graph 14)")

```

#### **CONCLUSION FROM GRAPH 12, GRAPH 13 AND GRAPH 14**
- Trend for an individual may be identified one at a time to find the relation between steps and heart rate.
- Also,more detailed data about the heart rate is needed to understand the trend

## -----------------------------------------------------------------------------------------
### RELATION BETWEEN THE AVERAGE ACTIVITY MINUTES AND HEARTRATE


```{r echo=FALSE}

#RELATION BETWEEN AVERAGE ACTIVITY MINUTES AND HEARTRATE
  #1  Average Heart Rate
ggplot(MergeActivitySleepHeartRate)+
  geom_point(mapping = aes(x=AvgActiveMin,y=AverageHeartRate,color=Id))+
labs(title = "Relation b/w Average Activity Minutes and Average Heart Rate for each user",caption = "(Graph 15)")
  #2  Maximum Heart Rate
ggplot(MergeActivitySleepHeartRate)+
  geom_point(mapping = aes(x=AvgActiveMin,y=MaxHeartRate,color=Id))+
labs(title = "Relation b/w Average Activity Minutes and Maximum Heart Rate for each user",caption = "(Graph 16)")
  #3  Minimum Heart Rate
ggplot(MergeActivitySleepHeartRate)+
  geom_point(mapping = aes(x=AvgActiveMin,y=MinHeartRate,color=Id))+
labs(title = "Relation b/w Average Activity Minutes and Minimum Heart Rate for each user",caption = "(Graph 17)")

```


#### **CONCLUSION FROM GRAPH 15, GRAPH 16 AND GRAPH 17**
- To understand the relation between the heart rate and activity minutes the minimum and maximum heartrate at the time of activity should be studied.

## -----------------------------------------------------------------------------------------
### RELATION BETWEEN THE CALORIES AND HEARTRATE

```{r echo=FALSE}

#RELATION BETWEEN CALORIES AND HEARTRATE
#1  Average Heart Rate
ggplot(MergeActivitySleepHeartRate)+
  geom_point(mapping = aes(x=Calories,y=AverageHeartRate,color=Id))+
labs(title = "Relation b/w Calories and Average Heart Rate for each user",caption = "(Graph 18)")

#2  Maximum Heart Rate
ggplot(MergeActivitySleepHeartRate)+
  geom_point(mapping = aes(x=Calories,y=MaxHeartRate,color=Id))+
  labs(title = "Relation b/w Calories and Maximum Heart Rate for each user",caption = "(Graph 19)")

#3  Minimum Heart Rate
ggplot(MergeActivitySleepHeartRate)+
  geom_point(mapping = aes(x=Calories,y=MinHeartRate,color=Id))+
  labs(title = "Relation b/w Calories and Minimum Heart Rate for each user",caption = "(Graph 20)")

```

#### **CONCLUSION FROM GRAPH 18, GRAPH 19 AND GRAPH 20**
- The relation between calories and heart rate seem to linear.
- More calories are burnt when heart rate increases.

## -----------------------------------------------------------------------------------------
### RELATION BETWEEN THE SLEEP AND HEARTRATE

```{r echo=FALSE}

#RELATION BETWEEN SLEEP AND HEARTRATE
#1  Average Heart Rate
ggplot(MergeActivitySleepHeartRate)+
  geom_point(mapping = aes(x=TotalMinutesAsleep,y=AverageHeartRate,color=Id))+
  labs(title = "Relation b/w Total Sleep minutes and Average Heart Rate for each user",caption = "(Graph 21)")

#2  Maximum Heart Rate
ggplot(MergeActivitySleepHeartRate)+
  geom_point(mapping = aes(x=TotalMinutesAsleep,y=MaxHeartRate,color=Id))+
  labs(title = "Relation b/w Total Sleep minutes and Maximum Heart Rate for each user",caption = "(Graph 22)")

#3  Minimum Heart Rate
ggplot(MergeActivitySleepHeartRate)+
  geom_point(mapping = aes(x=TotalMinutesAsleep,y=MinHeartRate,color=Id))+
labs(title = "Relation b/w Total Sleep minutes and Minimum Heart Rate for each user",caption = "(Graph 23)")
```

#### **CONCLUSION FROM GRAPH 21, GRAPH 22 AND GRAPH 23**
- More detailed data is needed about heartrate.

## -----------------------------------------------------------------------------------------
### RELATION BETWEEN THE INDIVIDUAL'S AVERAGE HEARTRATE AND WEIGHT


```{r echo=FALSE}
#RELATION BETWEEN AVERAGE HEARTRATE AND WEIGHT
ggplot(MergeActivitySleepHeartRateWeight)+
  geom_point(mapping = aes(x=WeightKg,y=AverageHeartRate,color=Id))+
  labs(title = "Relation b/w Weight in kg and Average Heart Rate for each user",caption = "(Graph 24)")
```

#### **CONCLUSION FROM GRAPH 24**
- More detailed data is needed about both heartrate and weight.


## -----------------------------------------------------------------------------------------
### RELATION BETWEEN THE SLEEP AND WEIGHT

```{r echo=FALSE}
#RELATION BETWEEN SLEEP AND WEIGHT
ggplot(MergeActivitySleepHeartRateWeight)+
  geom_point(mapping = aes(x=WeightKg,y=TotalMinutesAsleep,color=Id))+
  labs(title = "Relation b/w Weight in kg and Total Minutes Asleep for each user",caption = "(Graph 25)")
```


#### **CONCLUSION FROM GRAPH 25**
- More detailed data is needed about weight.


## -----------------------------------------------------------------------------------------
### RELATION BETWEEN THE SLEEP AND BMI

```{r echo=FALSE}
#RELATION BETWEEN SLEEP AND BMI
ggplot(MergeActivitySleepHeartRateWeight)+
  geom_point(mapping = aes(x=BMI,y=TotalMinutesAsleep,color=Id))+
  labs(title = "Relation b/w BMI and Total Minutes Asleep for each user",caption = "(Graph 26)")

#==========================================================================================
```

#### **CONCLUSION FROM GRAPH 26**
- More detailed data is needed about weight.

## -----------------------------------------------------------------------------------------

#### **RECOMMENDATIONS**
- Heart rate details are logged for only few users. More details could be collected to understand the trend.
- Weight data is also limited. Features could be added to track Fat and body measurements on the smart device itself.
- Logging weight with selection on smart device itself could be easy for users to make manual entry rather than going into the application.
- Features to track water intake easily with just one press on device could be added

## -----------------------------------------------------------------------------------------
## =======================================================================
